Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-08-20 manuscript.Rnw
Fixed table references

\renewcommand{\baselinestretch}{1.6}   % line spacing
\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
% \usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lineno}
\geometry{verbose,tmargin=2.4cm,bmargin=2.4cm,lmargin=2.4cm,rmargin=2.4cm}
\usepackage{natbib}

@@

We found that decomposing LMF and LAR as shown in Eq. \ref{eq:M1} was very useful to study climate effects because the second term ($A_S/M_T$) absorbs nearly all of the size related variation in LMF (Tables \ref{table/varpartB}, Table S1 and Fig. S3) and is otherwise fairly conserved across PFTs (Fig.~\ref{fig:fighist}). As shown in Fig.~\ref{fig:figlmfspecies}b and Fig.~\ref{fig:figlmf}d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Tables \ref{table/varpartB} and Table S1). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig.~\ref{fig:figclimate}). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table \ref{table/varpartB}). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig.~\ref{fig:figclimate}b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table S1).

@@

We found that climate did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for this difference is that our measurements were taken on individual plants, whereas in  \citet{reich2014} they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig.~\ref{fig:fighist}), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Tables \ref{table/varpartB} and Table S1).

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS2.pdf}
    \caption{Raw data for leaf biomass ($M_F$) and total above-ground woody biomass ($M_S$) for each of the PFTs, as a function of total plant height ($H$). Each point is an individual plant. Sample sizes are listed in Table~\ref{table:counttable}}
    \label{fig:figrawdata}
\end{figure}

@@

\textbf{Table S1} Explained variance by plant functional type (PFT), plant height (H), mean annual temperature (MAT) and mean annual precipitation (MAP), for five whole-plant variables.\\
\textbf{Table S5} Number of observations used in the analysis of leaf area ratio (LAR).\\
\textbf{Fig. S1} Global coverage of the climate space by the dataset, labelled by vegetation type.\\
\textbf{Fig. S2} Leaf area ratio ($A_F/M_T$) by PFT.\\
\textbf{Fig. S3} Relationship between above-ground biomass and basal stem area.\\
\textbf{Notes S1} Modelling of biomass partitioning in global vegetation models (GVMs)

>>>>

\renewcommand{\baselinestretch}{1.6}   % line spacing
\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
% \usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{lineno}
\geometry{verbose,tmargin=2.4cm,bmargin=2.4cm,lmargin=2.4cm,rmargin=2.4cm}
\usepackage{natbib}

@@

We found that decomposing LMF and LAR as shown in Eq. \ref{eq:M1} was very useful to study climate effects because the second term ($A_S/M_T$) absorbs nearly all of the size related variation in LMF (Table 2 and Fig. S3) and is otherwise fairly conserved across PFTs (Fig.~\ref{fig:fighist}). As shown in Fig.~\ref{fig:figlmfspecies}b and Fig.~\ref{fig:figlmf}d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Table 2). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig.~\ref{fig:figclimate}). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table 2). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig.~\ref{fig:figclimate}b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table 2).

@@

We found that climate did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for this difference is that our measurements were taken on individual plants, whereas in  \citet{reich2014} they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig.~\ref{fig:fighist}), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Table 2).

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS2.pdf}
    \caption{Raw data for leaf biomass ($M_F$) and total above-ground woody biomass ($M_S$) for each of the PFTs, as a function of total plant height ($H$). Each point is an individual plant. Sample sizes are listed in Table 1}
    \label{fig:figrawdata}
\end{figure}

@@

\textbf{Fig. S1} Global coverage of the climate space by the dataset, labelled by vegetation type.\\
\textbf{Fig. S2} Leaf area ratio ($A_F/M_T$) by PFT.\\
\textbf{Fig. S3} Relationship between above-ground biomass and basal stem area.\\
\textbf{Notes S1} Modelling of biomass partitioning in global vegetation models (GVMs)
